A Graph Theory Based Approach to MDAO Problem Formulation

### 2013-05-08 mcg_fpg.tex
Major update

\section{Building MCG and FPG}
    \label{s:building graphs}
\subsection{Maximal Connectivity Graph}
    \label{ss:MCG}
    To construct the maximal connectivity graph, we assume that a set of codes, global inputs, and objectives and constraints (collectively called global outputs) are provided. 
    The codes are represented by analysis block graphs ((ABGs?)) $A_i=(V_{A_i},E_{A_i}), \ i\in I$, where $m$ is the number of codes and $I=\{1,2,\ldots,m\}$. 
    The global inputs are represented as a set of variable nodes $V_\txt{in}$, and the global outputs are represented by a set of variable nodes $V_\txt{out}$. 
    We assume that $V_\txt{in}$, $V_\txt{out}$, and each $A_i$ are given, and that any potential connection between variables is given in the form of connection edges in the set $C_M$. 
    Then we may construct the maximal connectivity graph $M=(V_M,E_M)$ as
    \begin{IEEEeqnarray*}{rCl}
    V_M & = & V_\txt{in} \cup V_\txt{out} \cup \left( \bigcup_{i \in I} V_{A_i} \right), \\
    E_M & = & C_M \cup \left( \bigcup_{i \in I} E_{A_i} \right),
    \end{IEEEeqnarray*}
    The MCG $M$ is uniquely determined by the given set of analysis blocks, the required outputs, and the given global inputs. 
    In the cases where the set of global inputs is not known a priori, the process of obtaining the FPG will reveal the required inputs, as discussed subsequently.

\subsection{Fundamental Problem Graph}
    \label{ss:FPG}
    We now define the fundamental problem graph $F=(V_F,E_F)$ as a directed graph meeting the following conditions, which are explained subsequently,
    \begin{enumerate}
    \item[(1)] $F \subset M$ and $V_\txt{out} \subset V_F$
    \item[(2)] $\displaystyle{\forall i \in I, \txt{ if } F \cap A_i \neq \emptyset \txt{ then } A_i \subset F}$
    \item[(3)] $\displaystyle{\forall v \in V_F \txt{ with } t_\txt{node}(v) = \txt{`variable,'} \  \txt{deg}_l^-(v) \leq \txt{deg}^-(v) \leq \txt{deg}_u^-(v)}$
    \item[(4)] $\displaystyle{\forall v \in V_F \txt{ there exists a reverse path } P \subset R_F \txt{ from } x \txt{ to } v \txt{ with } x\in V_\txt{out}}$
    \end{enumerate}
    %The set $I_F$ is an index set containing the indices of the analysis blocks in $F$ and it may only include the analysis blocks $A_i$ that meet requirements (2) and (6). 
    %Requirement (2) stipulates that each analysis blocks in $F$ must have at least one local output that is being used. 
    %Requirement (3) stipulates that only the global inputs that are being used should be included. 
    %Requirements (4) and (5) provide the construction of the sets of nodes and edges, respectively. 
    Requirement (1) asserts that only the nodes and edges provided by the maximal connectivity graph can be used in the fundamental problem graph and that every global output must be included. 
    Requirement (2) requires that each analysis block must be entirely present or entirely absent. 
    Requirement (3) stipulates that the number of edges directed into each variable node must be within the lower and upper in-degree limits; 
    if $\txt{deg}^-(v) < \txt{deg}_l^-(v)$ the node is a \emph{hole}, and if $\txt{deg}^-(v) > \txt{deg}_u^-(v)$ the node is a \emph{collision}.
    Lastly, requirement (4) ensures that only the nodes that are being used are included in the FPG by requiring that for every node a reverse path exists from at least one global output to that node. 
    The reverse graph $R_F$ is obtained from $F$ by simply switching the orientation of every edge in $E_F$.

@@

\subsection{Obtaining the Fundamental Problem Graph}
    \label{ss:obtaining FPG}
    In general, there may be multiple different graphs that satisfy the FPG conditions in Sec.~\ref{ss:FPG}, though there may be none at all. 
    Here, we describe a process for obtaining an FPG by starting with the MCG and disconnecting connection edges until the FPG conditions are met. 
    Then the problem is reduced to deciding which connection edges to remove.

((describe pruning as a subroutine enforcing step 4 and then step 2))

    \begin{description}
    \item[Subroutine: Pruning] 
        Given a digraph $G_1 = (V_1,E_1)$, let the subroutine denoted as $P_\txt{prune}$ operate on $G_1$ to produce a digraph $G_2 = (V_2,E_2)$ that satisfies requirements (2) 
        and (4), and write $G_2 = P_\txt{prune} ( G_1 )$. 
        This is accomplished by first creating the reverse graph of $G_1$, $R = (V_1,E_R)$. Then adding a new node $x$ to $V_R$ and adding edges directed from this node to each of the global outputs, i.e.,
    \begin{equation}
        \forall y \in V_\txt{out}, (x,y) \in E_R
    \end{equation}
        The set of nodes that may be reached from the global outputs is constructed as
    \begin{equation}
        U = \{ n \in V_1 \st \exists P \txt{ a path from $x$ to $n$}, \ P \subset R \}.
        \end{equation}
        The list of analysis blocks with at least one node that may be reach is constructed as
        \begin{equation}
        I_1 = \{ i \in I \st V_{A_i} \cap U \neq \emptyset \}
        \end{equation}
        The set of nodes to remove is
        \begin{equation}
        N = \{ v \in V_1 \st v \in V_{A_i} \txt{ for } i \in I_1, \txt{ or } v \in V_\txt{in}\setminus U \},
        \end{equation}
        which means that $N$ is composed of unused analysis block nodes and unused global input nodes. Then the new set of nodes is created as
        \begin{equation}
        V_2 = V_1 \setminus N,
        \end{equation}
        and edges involving the removed nodes are also deleted
        \begin{equation}
        E_2 = E_1 \setminus \{ (a,b) \st a \in N \txt{ or } b \in N  \}.
        \end{equation}
        The set of connection edges can be obtained as
        \begin{equation}
        C_2 = \{(x,y) \in V_{2} \st \sim(x \in V_{A_i} \txt{ and } y \in V_{A_i} \txt{ for some } i \in I_1)\},
        \end{equation}
        which means $C_2 \cap V_{A_i} = \emptyset  \ \forall i \in I_1$, and $E_2 = C_2 \cup \left( \bigcup_{i \in I_1} E_{A_i} \right)$.

    \item[Step 1: Holes] 
        The first step is to detect holes and disconnect the connection edges following them. 
        These connection edges are removed because they represent variables which cannot be determined because the analysis function does not have adequate inputs.

        To start, the an initial graph is created as a pruned copy of the MCG
        \begin{equation}
        F_0 = P_\txt{prune}(M),
        \end{equation}
        where $F_0 = (V_{F,0},E_{F,0})$, and $C_{F,0}$ is also obtained as described. 
        The set of variable nodes which are holes is created as
        \begin{equation}
        H = \{v \in V_{F,0} \st t_\txt{node}(v) = \txt{`variable'} \txt{ and } \txt{deg}^-(v) < \txt{deg}_l^-(v) \},
        \end{equation}
        which is the set of variable nodes with fewer incoming edges than are allowed by the lower in-degree limit.
        Then the updated set of edges is created by removing the edges preceding or succeeding the analysis block:
        \begin{equation}
        C_{F,1} = C_{F,0} \setminus \{(x,y) \in C_{F,0} \st x \in V_{A_i} \txt{ or } y \in V_{A_i}, \txt{ and } H\cap V_{A_i} \neq \emptyset \},
    \end{equation}
    %deg-v is now recaculated for the new C. 
    and this step is demonstrated by Fig.~\ref{f:hole}. Because removing these edges can create new holes, this step must be repeated until no more holes are found.
        If this step every finds a global output to be a hole, meaning $V_\txt{out} \cap H \neq \emptyset$, then the algorithm stops because an FPG cannot be obtained. 
        Otherwise, it is guaranteed that an FPG can be obtained.
    \begin{figure}[htb!]
        \begin{center}
        \includegraphics[width=3in]{images/analysis_block_hole}
        \end{center}
        \vspace{-20pt}
    \caption{Example variable node indicating a hole.}
    \label{f:hole}
    \end{figure}

    \item[Step 2: Collisions] 
        The second step is to detect collisions and to then disconnect enough connection edges so that the collisions are resolved. 
        The set of variable nodes which contain collisions is created as
    \begin{equation}
    S_\txt{nodes} = \{v \in V_M \st t_\txt{node}(v) = \txt{`variable'} \txt{ and } \txt{deg}^-(v) > \txt{deg}_u^-(v) \}.
    \end{equation}
    For each collision node we can construct a set contained the edges directed in. The set containing all of these sets is constructed as
    \begin{equation}
        S_\txt{edges} = \big \{ \{(x,y) \in E_M\} \ \big| \ y \in S_\txt{nodes} \big \}
    \end{equation}
    Let $J=\{1,2,\ldots,|S_\txt{edges}|\}$ be an indexing set for $S_\txt{edges}$ such that each $S_{\txt{edges},j}$ corresponds to a set in $S_\txt{edges}$ for $j \in J$. 
    An example collision is shown in Fig.~\ref{f:collision} to indicate the definition of $S_{\txt{edges},j}$. 
    We can assume that $J$ also indexes $S_\txt{nodes}$ because there is a one--to--one correspondence between the elements in $S_\txt{nodes}$ and the elements in $S_\txt{edges}$ (which are sets). 
    Then we may construct sets of edges as
    \begin{equation}
    B_j = \big \{e_k \in S_{\txt{edges},j} \st k \in \{1,2,\ldots,K\} \txt{ with } K \leq \txt{deg}_u^-(v_j) \big \}, \ j \in J,
    \end{equation}
        which means that each set $B_j$ is constructed from the set $S_{\txt{edges},j}$ by taking only as many edges as are allowed by the upper in--degree limit of $v_j$. 
        The construction of each $B_j$ corresponds to making a decision on which edges to include and which edges not to include. Then let
    \begin{equation}
    C_{F,2} = \{ e \in C_{F,1} \st e \in B_j \txt{ for some } j \in J\},
    \end{equation}
    \begin{equation}
    E_{F,2} = E_M \setminus (C_M \setminus C_{F,2})
    \end{equation}
    and
    \begin{equation}
    F_2 = (V_M,E_{F,2})
    \end{equation}
    \begin{figure}[htb!]
        \begin{center}
        \includegraphics[width=3in]{images/analysis_block_collision}
        \end{center}
        \vspace{-20pt}
    \caption{Example variable node indicating a collision.}
    \label{f:collision}
    \end{figure}

    \item[Step 3: Finalize] 
        The third and final step is to prune the graph one last time
    \begin{equation}
        F = P_\txt{prune}(F_2),
    \end{equation}
        and $F = (V_F,E_F)$.
        The indices of the used analysis blocks can then be found as
    \begin{equation}
        I_F = \{ i \in I \st A_i \cap F \neq \emptyset \}
    \end{equation}
%       \begin{
%       \begin{equation}
%       I_F =  \{i \in I \ | \ \exists v \in V_{A_i} \txt{ such that }  t_\txt{node}(E_{F,2}^{-1}(v))=\txt{`model'} \txt{ and } \txt{deg}^+(v) > 0 \},
%       \end{equation}
%       where the degree is calculated with respect to $F_2$. 
%       This set excludes both the analysis blocks that were unused in the original MCG and those that became unused in step 1 or step 2.
%
%       The set of used global inputs is constructed as
%       \begin{equation}
%       V_{\txt{in},F} = \big\{v \in V_\txt{in} \ \big| \ |\{(x,y) \in E_{F,2}(v) \st y \in V_{A_i} \txt{ for } i \in I_F  \}| > 1 \big\},
%       \end{equation}
%       which requires that at least two edges directed out of the nodes in $V_{\txt{in},F}$ must be to the analysis blocks indexed by $I_F$.
%
%       Finally, the set of nodes and edges describing a fundamental problem graph is given by
%       \begin{equation}
%       V_F = V_{\txt{in},F} \cup V_\txt{out} \cup \left( \bigcup_{i \in I_F} V_{A_i} \right),
%       \end{equation}
%       \begin{equation}
%       C_{F} = \{ (x,y) \in C_{F,2} \st x,y \in V_F\},
%       \end{equation}
%       \begin{equation}
%       E_F = C_F \cup \left( \bigcup_{i \in I_F} E_{A_i} \right),
%       \end{equation}
%       \begin{equation}
%       F = (V_F,E_F).
%       \end{equation}

@@

This process will always provide an FPG if one exists. If an FPG does not exist, this process will reveal the limiting factors preventing a valid formulation.

>>>>

\section{The MCG and FPG}
    \label{s:building graphs}
We now discuss how the graph-based syntax defined in Sec.~\ref{s:syntax definition} is used to represent an MDAO problem in its entirety. This discussion is based on the MCG and the FPG.

\subsection{Maximal Connectivity Graph}
    \label{ss:MCG}
For a specified MDAO problem, the maximal connectivity graph represents every analysis tool, objective, constraint, and global input being considered, as well as every possible interconnection among them.
The definition of the MCG is given via construction.
    To construct the maximal connectivity graph, we presume that a set of analysis tools, global inputs, objectives, and constraints are provided. 
    Each of the $m$ analysis codes correspond to an index $i\in I_A$, $I_A=\{1,2,\ldots,m\}$, and are represented by an analysis block graph $G_{A_i}=(V_{A_i},E_{A_i})$.
    Each of the $n$ expression blocks correspond to an index $i\in I_E$, $I_E=\{1,2,\ldots,n\}$, and are represented by an expression block graph $G_{E_i}=(V_{E_i},E_{E_i})$.
    Finally, the global inputs are represented as a set of variable nodes $V_\txt{in}$.
    We presume that $V_\txt{in}$, each $G_{A,i}$, and each $G_{E,i}$ are given, and that any potential connection between variables is specified in the form of connection edges in the set $E_{M,\txt{con}}$. One method for defining these connections is to use a consistant variable naming convention.
    We may then construct the maximal connectivity graph $M=(V_M,E_M)$ as
    \begin{IEEEeqnarray*}{rCl}
    V_M & = & V_\txt{in} \cup \left( \bigcup_{i \in I_A} V_{A_i} \right) \cup \left( \bigcup_{i \in I_E} V_{E_i} \right), \\
    E_M & = & E_{M,\txt{con}} \cup \left( \bigcup_{i \in I_A} E_{A_i} \right)  \cup \left( \bigcup_{i \in I_E} E_{E_i} \right).
    \end{IEEEeqnarray*}
    The MCG $M$ is uniquely determined by the given set of analysis blocks, global inputs, and expression blocks. 
    In the cases where the set of global inputs is not known a priori, the process of obtaining the FPG will reveal the required inputs, as discussed subsequently.

The nodes and edges in $M$ can be partioned in sets according to their type:
\begin{equation}
V_M = V_{M,\txt{var}} \cup V_{M,\txt{mod}}, \ V_{M,\txt{var}} \cap V_{M,\txt{mod}} = \emptyset,
\end{equation}
where $V_{M,\txt{var}}$ and $V_{M,\txt{mod}}$ are the sets of variable nodes and model nodes, respecively;
\begin{equation}
E_M = E_{M,\txt{con}} \cup E_{M,\txt{mod}}, \ E_{M,\txt{con}} \cap E_{M,\txt{mod}} = \emptyset,
\end{equation}
where $ E_{M,\txt{con}}$ and $E_{M,\txt{mod}}$ are the sets of connection edges and model edges, respectively. These sets will be referenced in the process for obtaining an FPG.

\subsection{Fundamental Problem Graph}
    \label{ss:FPG}
    We now define the fundamental problem graph $F=(V_F,E_F)$ as a directed graph meeting the following conditions:
    \begin{enumerate}
    \item[(1)] $F \subset M$ and $G_{E_i} \subset F  \ \forall i \in I_E$
%    \item[(1)] $F \subset M$ and $V_\txt{out} \subset V_F$
    \item[(2)] $\displaystyle{\forall i \in I_A, \txt{ if } F \cap G_{A_i} \neq \emptyset \txt{ then } G_{A_i} \subset F}$
%    \item[(2)] $\displaystyle{\forall i \in I, \txt{ if } F \cap A_i \neq \emptyset \txt{ then } A_i \subset F}$
    \item[(3)] $\displaystyle{\forall v \in V_F \txt{ with }v \in V_{M,\txt{var}} \txt{ and }  \txt{deg}_l^-(v) \leq \txt{deg}^-(v) \leq \txt{deg}_u^-(v)}$
    \item[(4)] $\displaystyle{\forall v \in V_F \txt{ there exists a reverse path } P \subset R_F \txt{ from } x \txt{ to } v \txt{ with } x\in V_{E_i}}$
%    \item[(4)] $\displaystyle{\forall v \in V_F \txt{ there exists a reverse path } P \subset R_F \txt{ from } x \txt{ to } v \txt{ with } x\in V_\txt{out}}$
    \end{enumerate}
    %The set $I_F$ is an index set containing the indices of the analysis blocks in $F$ and it may only include the analysis blocks $A_i$ that meet requirements (2) and (6). 
    %Requirement (2) stipulates that each analysis blocks in $F$ must have at least one local output that is being used. 
    %Requirement (3) stipulates that only the global inputs that are being used should be included. 
    %Requirements (4) and (5) provide the construction of the sets of nodes and edges, respectively. 
    Condition (1) asserts that only the nodes and edges provided by the maximal connectivity graph can be used in the fundamental problem graph and that every expression block must be included. 
    Condition (2) requires that analysis blocks must be included or excluded in their entirety. 
    Condition (3) stipulates that the number of edges directed into each variable node must be within the lower and upper in-degree limits; 
    if $\txt{deg}^-(v) < \txt{deg}_l^-(v)$ the node is the location of a \emph{hole}, and if $\txt{deg}^-(v) > \txt{deg}_u^-(v)$ the node is the location of a \emph{collision}.
    Lastly, condition (4) ensures that only the nodes that are being used are included in the FPG by requiring that for every node a reverse path exists from at least one expression block to that node. 
    The reverse graph $R_F$ is obtained from $F$ by simply switching the orientation of every edge in $E_F$.

@@

\subsection{Algorithm for obtaining the Fundamental Problem Graph from the MCG}
    \label{ss:obtaining FPG}
    In general, there may be zero, one, or many graphs that satisfy the FPG conditions in Sec.~\ref{ss:FPG}.
    Here, we describe an algorithm for obtaining an FPG by starting with the MCG and disconnecting connection edges until the FPG conditions are met. 
    With this approach, the problem is reduced to deciding which connection edges to remove. First, we describe a pruning subroutine used within the algorithm, and next we describe the algorithm itself.

    \begin{description}
    \item[\bf{Subroutine: Pruning}] 
        Given a digraph $G_1 = (V_1,E_1)$, let the subroutine denoted as $P_\txt{prune}$ operate on $G_1$ to produce a new digraph $G_2$ as
	\begin{equation}
	G_2 = P_\txt{prune}(G_1) = (V_2,E_2),
	\end{equation} 
	where $G_2$ now satisfies requirements (2) and (4) in Sec.~\ref{ss:FPG}.
        Pruning is accomplished by first creating the reverse graph of $G_1$, $R = (V_1,E_R)$, where
	\begin{equation}
	E_R = \{ (x,y) \st (y,x) \in E_1\}.
	\end{equation}
	Next, a new node $b$ is added to $V_1$ and edges directed from this node to each of the expression blocks are added:
    \begin{equation}
        \forall i \in I_E,\forall v \in V_{E_i}, \txt{ if } \txt{deg}^+(v)=0 \txt{ then } (b,v) \in E_R
    \end{equation}
        The set of nodes that may be reached from $b$ is constructed as
    \begin{equation}
        U = \{ v \in V_1 \st \exists P \txt{ a path from $b$ to $v$}, \ P \subset R \}.
        \end{equation}
	Because $b$ is directed into only the expression blocks, any path from $b$ necessarily provides a path from at least one expression block, which means the node is being used in the problem formulation.

        The list of analysis blocks with at least one node in $U$ is constructed as
        \begin{equation}
        I_2 = \{ i \in I_A \st V_{A_i} \cap U \neq \emptyset \}
        \end{equation}
        Any analysis block not in $I_2$ should be removed from the graph because none of its outputs contribute to the problem. Global inputs that are not being used are also removed. The set of nodes to remove is then
        \begin{equation}
        N = (V_\txt{in}\setminus U )  \cup \left( \bigcup_{i \notin I_2} V_{A_i} \right),
        \end{equation}
and the new set of nodes is created as
        \begin{equation}
        V_2 = V_1 \setminus N.
        \end{equation}
        Edges involving the removed nodes are also deleted as
        \begin{equation}
        E_2 = E_1 \setminus \{ (x,y) \st x \in N \txt{ or } y \in N  \}.
        \end{equation}
        The set of connection edges can be extracted by considering only edges whos endpoints are not in the same analysis block:
        \begin{equation}
        E_{2,\txt{con}} = \{(x,y) \in V_{2} \st \sim(x \in V_{A_i} \txt{ and } y \in V_{A_i} \txt{ for some } i \in I_2)\}.
        \end{equation}
        %which means $C_2 \cap V_{A_i} = \emptyset  \ \forall i \in I_1$, and $E_2 = C_2 \cup \left( \bigcup_{i \in I_1} E_{A_i} \right)$. 
\end{description}

\noindent The algorithm for obtaining an FPG can now be described as follows:
\begin{description}
    \item[\bf{Step 1: Holes}] 
        The first step is to detect holes and disconnect the first set of connection edges downstream of them, as indicated in Fig.~\ref{f:hole}. 
        These connection edges are removed because they represent output variables which cannot be determined because the analysis function does not have adequate inputs.

        To begin the process, an initial graph is created as a pruned copy of the MCG
        \begin{equation}
        F_0 = P_\txt{prune}(M),
        \end{equation}
        where $F_0 = (V_{F_0},E_{F_0})$, and $C_{F_0,\txt{con}}$ is also obtained as described. 
        The set of variable nodes which are holes is identified as
        \begin{equation}
        H = \{v \in V_{F_0} \st v \in V_{M,\txt{var}} \txt{ and } \txt{deg}^-(v) < \txt{deg}_l^-(v) \},
        \end{equation}
        which is the set of variable nodes with fewer incoming edges than are allowed by the lower indegree limit.
        The updated set of connection edges is then created by removing the edges preceding or succeeding the analysis block:
        \begin{equation}
        E_{F_1,\txt{con}} = E_{F_0,\txt{con}} \setminus \{(x,y) \in E_{F_0,\txt{con}} \st x \in V_{A_i} \txt{ or } y \in V_{A_i}, \txt{ and } H\cap V_{A_i} \neq \emptyset \}.
    \end{equation}
Because removing these edges can create new holes, this step must be repeated until no additional holes are found.
        If the hole identification step identifies a variable node in an expression block as hole, meaning $V_{E_i} \cap H \neq \emptyset$ for some $i \in I_E$, then the algorithm terminates because an FPG cannot be obtained.

If this step finishes without finding a hole in an expression block, it is guaranteed that an FPG can be obtained because $F_1$ now satisfies all four conditions from Sec.~\ref{ss:FPG} expect for (3), which requires there to be no holes or collisions in the graph. Since no more holes remain, and collisions can always be resolved without producing a hole, (see Table \ref{t:variable node classification}), then it is guaranteed than an FPG can be obtained.

    \begin{figure}[htb!]
        \begin{center}
        \includegraphics[width=3in]{images/analysis_block_hole}
        \end{center}
        \vspace{-20pt}
    \caption{Example variable node indicating a hole.}
    \label{f:hole}
    \end{figure}

    \item[\bf{Step 2: Collisions}] 
        The second step is to detect collisions and to disconnect precisely the number of connection edges required such that all collisions are resolved without introducting holes. 
        The set of variable nodes at which collisions occur is 
    \begin{equation}
    S_\txt{nodes} = \{v \in V_{M,\txt{var}} \st \txt{deg}^-(v) > \txt{deg}_u^-(v) \}.
    \end{equation}
    For each collision node we can construct a set containing the edges directed into the node. The set containing all of these sets is constructed as
    \begin{equation}
        S_\txt{edges} = \big \{ \{(x,y) \in E_M\} \ \big| \ y \in S_\txt{nodes} \big \}
    \end{equation}
    Let $J=\{1,2,\ldots,|S_\txt{edges}|\}$ be an indexing set for $S_\txt{edges}$ such that each $S_{\txt{edges},j}$ corresponds to a set in $S_\txt{edges}$ for $j \in J$. 
    An example collision is shown in Fig.~\ref{f:collision} to exemplify the definition of $S_{\txt{edges},j}$. 
    We can assume that $J$ also indexes $S_\txt{nodes}$ because there is a one--to--one correspondence between the elements in $S_\txt{nodes}$ and the elements in $S_\txt{edges}$ (which are sets). 
    We may then construct sets of edges as
    \begin{equation}
    B_j = \big \{e_k \in S_{\txt{edges},j} \st k \in \{1,2,\ldots,K\} \txt{ with } \txt{deg}_u^-(v_j) \leq K \leq \txt{deg}_u^-(v_j) \big \}, \ j \in J,
    \end{equation}
        which means that each set $B_j$ is constructed from the set $S_{\txt{edges},j}$ by taking only as many edges as are allowed by the upper and lower indegree limits of $v_j$.  
The construction of each $B_j$ corresponds to making a decision about which edges to include and which edges not to include. 
%An upper bound on the number of such decisions that need to be made is 
%\begin{equation}
%\prod_{j \in J} 	
%	\begin{pmatrix}
%		|S_{\txt{edges},j}| \\
%		\txt{deg}_u^-(v_j)
%	\end{pmatrix}
%\end{equation}
%where $(:)$ refers to standard $n$-choose-$k$ notation.
Let the new set of connection edges be
    \begin{equation}
    E_{F_2,\txt{con}} = \{ e \in E_{F_1,\txt{con}} \st e \in B_j \txt{ for some } j \in J\}.
    \end{equation}
The set of all edges is created by removing the connection edges not in $E_{F_2,\txt{con}}$:
    \begin{equation}
    E_{F,2} = E_{F_0} \setminus (E_{M,\txt{con}} \setminus E_{F_2,\txt{con}}),
    \end{equation}
    which gives
    \begin{equation}
    F_2 = (V_{F_0},E_{F,2}).
    \end{equation}
    \begin{figure}[htb!]
        \begin{center}
        \includegraphics[width=3in]{images/analysis_block_collision}
        \end{center}
        \vspace{-20pt}
    \caption{Example variable node indicating a collision.}
    \label{f:collision}
    \end{figure}

    \item[\bf{Step 3: Finalize}] 
        The third and final step is to prune the graph a final time to exclude any analysis blocks that became unneeded after the collisions were resolved:
    \begin{equation}
        F = P_\txt{prune}(F_2),
    \end{equation}
        with $F = (V_F,E_F)$.
        The indices of the analysis blocks implemented in the FPG can then be found as
    \begin{equation}
        I_F = \{ i \in I_A \st G_{A_i} \subset F \}.
    \end{equation}
%       \begin{
%       \begin{equation}
%       I_F =  \{i \in I \ | \ \exists v \in V_{A_i} \txt{ such that }  t_\txt{node}(E_{F,2}^{-1}(v))=\txt{`model'} \txt{ and } \txt{deg}^+(v) > 0 \},
%       \end{equation}
%       where the degree is calculated with respect to $F_2$. 
%       This set excludes both the analysis blocks that were unused in the original MCG and those that became unused in step 1 or step 2.
%
%       The set of used global inputs is constructed as
%       \begin{equation}
%       V_{\txt{in},F} = \big\{v \in V_\txt{in} \ \big| \ |\{(x,y) \in E_{F,2}(v) \st y \in V_{A_i} \txt{ for } i \in I_F  \}| > 1 \big\},
%       \end{equation}
%       which requires that at least two edges directed out of the nodes in $V_{\txt{in},F}$ must be to the analysis blocks indexed by $I_F$.
%
%       Finally, the set of nodes and edges describing a fundamental problem graph is given by
%       \begin{equation}
%       V_F = V_{\txt{in},F} \cup V_\txt{out} \cup \left( \bigcup_{i \in I_F} V_{A_i} \right),
%       \end{equation}
%       \begin{equation}
%       C_{F} = \{ (x,y) \in C_{F,2} \st x,y \in V_F\},
%       \end{equation}
%       \begin{equation}
%       E_F = C_F \cup \left( \bigcup_{i \in I_F} E_{A_i} \right),
%       \end{equation}
%       \begin{equation}
%       F = (V_F,E_F).
%       \end{equation}

@@

This algorithm will always provide an FPG if one exists. If an FPG does not exist, the limiting factors preventing a valid formulation are revealed.

\subsection{Suggested Process for Using the FPG Algorithm}
\label{ss:process}
Section \ref{ss:obtaining FPG} provided an algorithm for obtaining an FPG from the given MCG. However, if an FPG cannot be obtained, this algorithm can be applied as part of a larger process in which the designer changes the problem or the supplied analyses to attempt to obtain an FPG. The following steps detail the suggested procedure for obtaining an FPG:
\begin{enumerate}
\item[\bf{(A)}] Begin with a set of global inputs, analysis codes, objectives, and constraints.
\item[\bf{(B)}] Build the MCG as described in Sec.~\ref{ss:MCG}.
\item[\bf{(C)}] Set indegree limits for variable nodes representing local inputs as described in Sec.~\ref{s:indegree-outdegree} and Table \ref{t:variable node classification}
\item[\bf{(D)}] Run the FPG algorithm described in Sec.~\ref{ss:obtaining FPG}. If a valid FPG is unattainable:
		\begin{enumerate}
\item Change the MCG by adding analysis blocks and/or global inputs, which means identifying new analyses to include in a potential MDAO workflow.
\item Change indegree limits (see Table \ref{t:variable node classification}).
\end{enumerate}
\item[\bf{(E)}] Repeat from step (A) until an FPG is obtained.
\end{enumerate}

### 2013-05-08 problem_formulation.tex
updated per brian's suggestions
Major update

\documentclass[referee]{svjour3} % insert '[draft]' option to show overfull boxes

@@

\author{
  David Pate, %
  Justin Gray,%
  Dr. Brian German 
 }

 \institute { Justin Gray \at NASA Glenn Research Center, Mail Stop 5-11, 21000 Brookpark Rd Clevland OH 44135 %
  \and David Pate \at Gatech \and Dr. Brian German \at Gatech}

@@

\begin{abstract}
   blah blah blah
\end{abstract}

@@

\input{intro}
\input{graph_syntax}
\input{mcg_fpg}
\input{example_problem}

\section{Applications}

\section{Conclusions}

>>>>

%\documentclass[referee]{svjour3} % insert '[draft]' option to show overfull boxes
\documentclass[onecolumn]{svjour3} % insert '[draft]' option to show overfull boxes

@@

\author{
  David J. Pate, %
  Justin Gray, %
  Brian J. German 
 }

 \institute { Justin Gray 
	\at NASA Glenn Research Center, Mail Stop 5-11, 21000 
		Brookpark Rd Clevland OH 44135 %
	\and David J. Pate 
	\at Graduate Research Assistant, Georgia Institute of Technology, 270 Ferst Drive, Atlanta, GA, 30332, U.S.A.
	\and Brian German 
	\at Assistant Professor, Georgia Institute of Technology, 270 Ferst Drive, Atlanta, GA, 30332, U.S.A.}

@@

\begin{abstract}
The formulation of multidisciplinary design, analysis, and optimization (MDAO) problems has become increasingly complex as the number of analysis tools and design variables included in typical studies has grown.  This growth in the scale and scope of MDAO problems has been motivated by the need to incorporate additional design disciplines and to expand the parametric design space to enable the exploration of unconventional design concepts.  In this context, given a large set of disciplinary analysis tools, the problem of determining a feasible data flow between tools to produce a specified set of system-level outputs is combinatorically challenging.   The difficulty is compounded in multi-fidelity problems, which are of increasing interest to the MDAO community.  In this paper, we propose an approach for addressing this problem based on the formalism of graph theory.  The approach begins by constructing the maximal connectivity graph (MCG) describing all possible interconnections between a set of analysis tools. Graph operations are then conducted to reduce the MCG to a fundamental problem graph (FPG) that describes the connectivity of analysis tools needed to solve a specified system-level design problem. The FPG does not predispose a particular solution procedure; any relevant MDO solution architecture could be selected to implement the optimization.  The approach is applied to an example problem to formulate an FPG for a commercial aircraft MDAO study.
\end{abstract}

@@

\input{intro}
\input{graph_syntax}
\input{mcg_fpg}
\input{example_problem}
\input{conclusions}
